---
title: "Alcohol Consumption by Country"
author: "Joseph T. Powers"  
date: "2017-07-17"
output: github_document
always_allow_html: yes
---

# Overview
This script explores global norms of alcohol consumption. It's questions were driven by some uninformed hypotheses I held about how poverty and cold weather might drive higher alcohol consumption. 

## TL;DR 
People in wealthy countries drink more than people in poorer countries. People drink more in cold climates, but there is an interesting reversal in the wealthier countries of the northernmost latitudes.

```{r global options, include=FALSE}
knitr::opts_chunk$set(
  fig.path = 'figs/', fig.show = 'asis', dpi = 300, 
  include = TRUE, echo = FALSE, 
  warning = FALSE, message = FALSE, cache = FALSE
) 
```

```{r}
library(breadcrumbs)
library(kableExtra)
library(knitr)
library(tidyverse)
library(readr)
library(stringr)
library(forcats)
library(scales)
source_file_paths()
theme_set(theme_minimal())
```

```{r}
library(breadcrumbs)
library(tidyverse)
source_file_paths()

read_lines(file_consumption_raw, n_max = 5)
consumption <- read_csv(file_consumption_raw)

read_lines(file_longitude_latitude_raw, n_max = 5)
longlat <- read_csv(file_longitude_latitude_raw)

readLines(file_gdp_ppp_raw, n = 10)
ppp <- read_csv(file_gdp_ppp_raw, skip = 3)
```

```{r}
## clean up data ----
##############################################################################-

consumption <- consumption %>% rename(total = total_litres_of_pure_alcohol)

longlat <- 
  longlat %>% 
  rename(abbr_2 = country, country = name)

ppp <- 
  ppp %>% 
  select(country = `Country Name`, abbr = `Country Code`, ppp = `2016`)

# join the data 
drinks <- consumption %>% 
  left_join(longlat, by = "country") %>% 
  left_join(ppp, by = "country")
```